Between-dataset comparisons

### 2019-11-18 analysis/reports/cross_dataset.Rmd
fix some formatting
boxplots, proportion low and high plots

skewhists <- ggplot(data = all_di, aes(x = skew_percentile)) +
  geom_histogram(binwidth = 5, alpha = .3) +
  geom_histogram(binwidth = 5, data = filter(all_di, nsamples == 2500)) +
  theme_bw() +
  facet_grid(rows = vars(dat), cols = vars(singletons), scales = "free_y") +
  ggtitle("Skewness")
skewhists

simpsonhists <- ggplot(data = all_di, aes(x = simpson_percentile)) +
  geom_histogram(binwidth = 5, alpha = .3) +
  geom_histogram(binwidth = 5, data = filter(all_di, nsamples == 2500)) +
  theme_bw() +
  facet_grid(rows = vars(dat), cols = vars(singletons), scales = "free_y") +
  ggtitle("Simpson")
simpsonhists

@@

* FIA really struggles to get 2500 samples
* Portal is the *most* concentrated of all the datasets.
* Then MCDB and misc, then gentry, then bbs. FIA is hard to place but looks idiosyncratic.
* Curious about how variation in %ile maps on to S and N, *particularly* for Gentry and BBS. 
* Skewness is less extreme than Simpson (more density in the in-between). 
* Singletons doesn't appear to be terrifically impactful. From now on filtering to *unaltered* vectors.

```{r s and n mapping, fig.width = 12, fig.height = 10}

>>>>

skewhists <- ggplot(data = all_di, aes(x = skew_percentile)) +
  geom_histogram(binwidth = 5, alpha = .3, boundary = 0) +
  geom_histogram(binwidth = 5, data = filter(all_di, nsamples == 2500), boundary = 0) +
  theme_bw() +
  facet_grid(rows = vars(dat), cols = vars(singletons), scales = "free_y") +
  ggtitle("Skewness") +
  geom_vline(xintercept = 95, color = "red")
skewhists

simpsonhists <- ggplot(data = all_di, aes(x = simpson_percentile)) +
  geom_histogram(binwidth = 5, alpha = .3, boundary = 0) +
  geom_histogram(binwidth = 5, data = filter(all_di, nsamples == 2500), boundary = 0) +
  theme_bw() +
  facet_grid(rows = vars(dat), cols = vars(singletons), scales = "free_y") +
  ggtitle("Simpson") +
  geom_vline(xintercept = 5, color = "red")

simpsonhists

@@

* FIA really struggles to get 2500 samples
* Portal is the *most* concentrated of all the datasets.
* Then MCDB and misc, then gentry, then bbs. FIA is hard to place but looks idiosyncratic.
* Curious about how variation in %ile maps on to S and N, *particularly* for Gentry and BBS. 
* Skewness is less extreme than Simpson (more density in the in-between). 
* Singletons doesn't appear to be terrifically impactful. From now on filtering to *unaltered* vectors.
* I think it's best to filter to at least 2000 samples.

Some summary statistics:

```{r summary stats}

di_summary <- all_di %>%
  filter(nsamples >= 2000) %>%
  group_by(dat, singletons) %>%
  summarize(mean_skew = mean(skew_percentile, na.rm=T),
            sd_skew = sd(skew_percentile, na.rm = T),
            mean_simpson = mean(simpson_percentile, na.rm = T),
            sd_simpson = sd(simpson_percentile, na.rm = T),
            mean_nsamples = mean(nsamples),
            nsites = length(unique(site)),
            prop_skew_high = mean(skew_percentile >= 95),
            prop_simpson_low = mean(simpson_percentile <= 5),
            med_skew = median(skew_percentile, na.rm = T),
            med_simpson = median(simpson_percentile, na.rm = T),
            skew_25 = quantile(skew_percentile, probs = .25),
            simpson_75 = quantile(simpson_percentile, probs = .75)) %>%
  ungroup()

skew_boxplot <- ggplot(data = filter(all_di, nsamples >= 2000), aes(x = singletons, y = skew_percentile, group = singletons)) +
  # geom_point() +
  # geom_errorbar(aes(x = singletons, ymin = skew_25, ymax = med_skew)) +
  # geom_point(data = all_di, aes(x = singletons, y = skew_percentile), size = .5, alpha = .05) +
  # geom_label(aes(x = singletons, y = 10, label = signif(med_skew, digits = 4))) +
  geom_boxplot() +
  theme_bw() +
  facet_wrap(vars(dat)) +
  geom_hline(yintercept = 100, linetype = 2, color = "red") +
  ylim(0, 110) +
  ggtitle("Skewness, >= 2000 samples")
skew_boxplot

simpson_boxplot <- ggplot(data = filter(all_di, nsamples >= 2000), aes(x = singletons, y = simpson_percentile, group = singletons)) +
  geom_boxplot() +
  theme_bw() +
  facet_wrap(vars(dat)) +
  geom_hline(yintercept = 0, linetype = 2, color = "red") +
  ylim(-10, 100) +
  ggtitle("Simpson's, >= 2000 samples")
simpson_boxplot

print(select(di_summary, dat, singletons, mean_nsamples, nsites, prop_skew_high, prop_simpson_low))

skew_summary_plot <- ggplot(data = di_summary, aes(x = dat, y = prop_skew_high, color = singletons, size = as.factor(nsites))) +
  geom_point() +
  theme_bw() +
  ggtitle("Skew") +
  scale_color_viridis_d(end = .8) +
  ylim(0, 1) +
  theme(legend.position = "none")

simpson_summary_plot <- ggplot(data = di_summary, aes(x = dat, y = prop_simpson_low, color = singletons, size = as.factor(nsites))) +
  geom_point() +
  theme_bw() +
  ggtitle("Simpson") +
  scale_color_viridis_d(end = .8)+
  ylim(0, 1) +
  theme(legend.position = "none")

gridExtra::grid.arrange(grobs = list(skew_summary_plot, simpson_summary_plot), nrow = 1)

```{r s and n mapping, fig.width = 12, fig.height = 18}
